---
title: "Sparse Matrix in RcppArmadillo"
output:
  pdf_document: default
  html_document: default
date: '`r format(Sys.time(), "%F %T %z")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Sparse Matrix in RcppArmadillo

## Introduction

The following documentation is intended for RcppArmadillo sparse matrix users' convenience based on the documentation of library [Matrix](https://cran.r-project.org/web/packages/Matrix/Matrix.pdf) and [Armadillo](http://arma.sourceforge.net/docs.html#SpMat). The [Unofficial Rcpp API](https://github.com/binxiangni/rcpp-api/blob/master/rcpp-api-docs.Rmd) also helps during integration of this documentation. Suggestions and corrections are welcomed.  

## Overview

- [Sparse Matrix in "RcppArmadillo"](#smir)
    - [Sparse Matrix in "Matrix"](#smim)
    - [Sparse Matrix in "Armadillo"](#smia)
- [Member functions](#member-funcs)
    - [Element Access](#elem-access)
    - [Fundamental Arithmetic Operations](#operations)
    - [Submatrix Views](#sub-views)
    - [Diagonal Views](#diag-views)
    - [Savings & Loadings](#save-load)
    - [Element-wise Functions](#elem-wise-funcs)
    - [Scalar Functions](#scalar-funcs)
    - [Vector Valued Functions](#vec-funcs)
    - [Matrix Valued Functions](#mat-funcs)
    - [Generated Matrices](#generated-mats)
    - [Eigen & Singular Value Decomposition](#evd-svd)
    - [Solution of Sparse Linear Systems](#solve)
    - [Miscellaneous](#misc)

## Sparse Matrix in "RcppArmadillo" {#smir}
### Sparse Matrix in "Matrix" {#smim}
There are 31 types of sparse matrices in the package "Matrix" that can be directly used. But for now, only 12 of them are supported in RcppArmadillo: [dgCMatrix](#dgc), [dtCMatrix](#dtc), [dsCMatrix](#dsc), [dgTMatrix](#dgt), [dtTMatrix](#dtt), [dsTMatrix](#dst), [dgRMatrix](#dgr), [dtRMatrix](#dtr), [dsRMatrix](#dsr), [indMatrix](#ind), [pMatrix](#pm), and [ddiMatrix](#ddi). 

#### dgCMatrix {#dgc}
- Description: general column-oriented numeric sparse matrix. 
- Constructor    
    - `new("dgCMatrix", ...)`  
    - `Matrix(*, sparse = TRUE)`  
    - `sparseMatrix()`
- Coercion    
    - `as(*, "CsparseMatrix")`  
    - `as(*, "dgCMatrix")`  
- Examples   
```{r dgCMatrix}
(dgC <- new("dgCMatrix", i = as.integer(c(0, 0, 1)), 
            p = as.integer(c(0, 1, 3)), 
            x = as.double(c(1, 2, 1)), 
            Dim = as.integer(c(3, 2))))

# Output
# 3 x 2 sparse Matrix of class "dgCMatrix"
#         
# [1,] 1 2
# [2,] . 1
# [3,] . .

x <- matrix(c(1, 0, 0, 2, 1, 0), nrow = 3)
(SM <- Matrix(x, sparse = TRUE))

# Output
# 3 x 2 sparse Matrix of class "dgCMatrix"
#         
# [1,] 1 2
# [2,] . 1
# [3,] . .

i <- c(1,3:8); j <- c(2,9,6:10); x <- 7 * (1:7)
(A <- sparseMatrix(i, j, x = x))  

# Output
# 8 x 10 sparse Matrix of class "dgCMatrix"
#                              
# [1,] . 7 . . .  .  .  .  .  .
# [2,] . . . . .  .  .  .  .  .
# [3,] . . . . .  .  .  . 14  .
# [4,] . . . . . 21  .  .  .  .
# [5,] . . . . .  . 28  .  .  .
# [6,] . . . . .  .  . 35  .  .
# [7,] . . . . .  .  .  . 42  .
# [8,] . . . . .  .  .  .  . 49

m <- matrix(c(1, 0, 0, 2, 1, 0), 3, 2)
(dgC <- as(m, "CsparseMatrix"))

# Output
# 3 x 2 sparse Matrix of class "dgCMatrix"
#         
# [1,] 1 2
# [2,] . 1
# [3,] . .

(dgC <- as(m, "dgCMatrix"))

# Output
# 3 x 2 sparse Matrix of class "dgCMatrix"
#         
# [1,] 1 2
# [2,] . 1
# [3,] . .

```

#### dtCMatrix {#dtc}
- Description: triangular column-oriented numeric sparse matrix. 
- Constructor   
    - `new("dtCMatrix", ...)`    
    - `Matrix(*, sparse = TRUE)`
    - `sparseMatrix(*, triangular=TRUE)`
- Coercion  
    - `as(*, "triangularMatrix")`  
    - `as(*, "dtCMatrix")`  
- Examples  
```{r dtCMatrix}
(t5 <- new("dtCMatrix", Dim = c(5L, 5L), uplo = "L",
                  x = c(10, 1, 3, 10, 1, 10, 1, 10, 10),
                  i = c(0L, 2L, 4L, 1L, 3L,2L, 4L, 3L, 4L),
                  p = c(0L, 3L, 5L, 7:9)))

# Output
# 5 x 5 sparse Matrix of class "dtCMatrix"
#                    
# [1,] 10  .  .  .  .
# [2,]  . 10  .  .  .
# [3,]  1  . 10  .  .
# [4,]  .  1  . 10  .
# [5,]  3  .  1  . 10

x <- matrix(c(2:0, 1), nrow = 2)
(dtC <- Matrix(x, sparse = TRUE))

# Output
# 2 x 2 sparse Matrix of class "dtCMatrix"
#         
# [1,] 2 .
# [2,] 1 1

i <- c(1,3:8); j <- c(2,9,6:10); x <- 7 * (1:7)
(tA <- sparseMatrix(i, j, x = x, triangular= TRUE))

# Output
# 10 x 10 sparse Matrix of class "dtCMatrix"
#                               
#  [1,] . 7 . . .  .  .  .  .  .
#  [2,] . . . . .  .  .  .  .  .
#  [3,] . . . . .  .  .  . 14  .
#  [4,] . . . . . 21  .  .  .  .
#  [5,] . . . . .  . 28  .  .  .
#  [6,] . . . . .  .  . 35  .  .
#  [7,] . . . . .  .  .  . 42  .
#  [8,] . . . . .  .  .  .  . 49
#  [9,] . . . . .  .  .  .  .  .
# [10,] . . . . .  .  .  .  .  .

dgC <- as(x, "dgCMatrix")
(dtC <- as(dgC, "triangularMatrix"))
# Output
# 2 x 2 sparse Matrix of class "dtCMatrix"
#         
# [1,] 2 .
# [2,] 1 1

(dtC <- as(dgC, "dtCMatrix"))
# Output
# 2 x 2 sparse Matrix of class "dtCMatrix"
#         
# [1,] 2 .
# [2,] 1 1
```

#### dsCMatrix {#dsc}
- Description: symmetric column-oriented numeric sparse matrix. 
- Constructor  
    - `new("dsCMatrix", ...)`  
    - `Matrix(*, sparse = TRUE)`  
    - `sparseMatrix(*, symmetric = TRUE)`  
- Coercion  
    - `as(*, "symmetricMatrix")`   
    - `as(*, "dsCMatrix")`   
- Examples  
```{r dsCMatrix}
(dsCMatrix <- new("dsCMatrix", 
                  i = c(0L, 1L, 0L, 2L), 
                  p = c(0L, 1L, 2L, 4L), 
                  x = rep(5, 4), 
                  Dim = c(3L, 3L)))

# Output
# 3 x 3 sparse Matrix of class "dsCMatrix"
#           
# [1,] 5 . 5
# [2,] . 5 .
# [3,] 5 . 5

x <- matrix(c(rep(c(5, 0), 4), 5), nrow = 3)
(dsC <- Matrix(x, sparse = TRUE))

# Output
# 3 x 3 sparse Matrix of class "dsCMatrix"
#           
# [1,] 5 . 5
# [2,] . 5 .
# [3,] 5 . 5

i <- c(1,3:8); j <- c(2,9,6:10); x <- 7 * (1:7)
(sA <- sparseMatrix(i, j, x = x, symmetric = TRUE))

# Output
# 10 x 10 sparse Matrix of class "dsCMatrix"
#                                  
#  [1,] . 7  .  .  .  .  .  .  .  .
#  [2,] 7 .  .  .  .  .  .  .  .  .
#  [3,] . .  .  .  .  .  .  . 14  .
#  [4,] . .  .  .  . 21  .  .  .  .
#  [5,] . .  .  .  .  . 28  .  .  .
#  [6,] . .  . 21  .  .  . 35  .  .
#  [7,] . .  .  . 28  .  .  . 42  .
#  [8,] . .  .  .  . 35  .  .  . 49
#  [9,] . . 14  .  .  . 42  .  .  .
# [10,] . .  .  .  .  .  . 49  .  .

dgC <- as(x, "dgCMatrix")
(dsC <- as(dgC, "symmetricMatrix"))

# Output
# 3 x 3 sparse Matrix of class "dsCMatrix"
#           
# [1,] 5 . 5
# [2,] . 5 .
# [3,] 5 . 5

(dsC <- as(x, "dsCMatrix"))

# Output
# 3 x 3 sparse Matrix of class "dsCMatrix"
#           
# [1,] 5 . 5
# [2,] . 5 .
# [3,] 5 . 5
```

#### dgTMatrix {#dgt}
- Description: general numeric sparse matrix in triplet form. 
- Constructor    
    - `new("dgTMatrix", ...)`  
    - `sparseMatrix(*, giveCsparse=FALSE)` 
    - `spMatrix()`
- Coercion  
    - `as(*, "TsparseMatrix")`
    - `as(*, "dgTMatrix")`  
- Examples   
```{r dgTMatrix}
(T2 <- new("dgTMatrix",
          i = c(1L,1L,0L,3L,3L),
          j = c(2L,2L,4L,0L,0L), 
          x=10*1:5, Dim=4:5))

# Output
# 4 x 5 sparse Matrix of class "dgTMatrix"
#                  
# [1,]  . .  . . 30
# [2,]  . . 30 .  .
# [3,]  . .  . .  .
# [4,] 90 .  . .  .

(m <- sparseMatrix(i= 1:8,j=2:9,
                   x=c(0:2,3:-1),giveCsparse=FALSE))

# Output
# 8 x 9 sparse Matrix of class "dgTMatrix"
#                        
# [1,] . 0 . . . . . .  .
# [2,] . . 1 . . . . .  .
# [3,] . . . 2 . . . .  .
# [4,] . . . . 3 . . .  .
# [5,] . . . . . 2 . .  .
# [6,] . . . . . . 1 .  .
# [7,] . . . . . . . 0  .
# [8,] . . . . . . . . -1

(m <- spMatrix(8,9, i= 1:8, j=2:9, x = c(0:2,3:-1)))

# Output
# 8 x 9 sparse Matrix of class "dgTMatrix"
#                        
# [1,] . 0 . . . . . .  .
# [2,] . . 1 . . . . .  .
# [3,] . . . 2 . . . .  .
# [4,] . . . . 3 . . .  .
# [5,] . . . . . 2 . .  .
# [6,] . . . . . . 1 .  .
# [7,] . . . . . . . 0  .
# [8,] . . . . . . . . -1

x <- matrix(c(1, 0, 0, 2, 1, 0), nrow = 3)
(dgT <- as(x, "TsparseMatrix"))

# Output
# 3 x 2 sparse Matrix of class "dgTMatrix"
#         
# [1,] 1 2
# [2,] . 1
# [3,] . .

(dgT <- as(x, "dgTMatrix"))
# Output
# 3 x 2 sparse Matrix of class "dgTMatrix"
#         
# [1,] 1 2
# [2,] . 1
# [3,] . .
```

#### dtTMatrix {#dtt}
- Description: triangular numeric sparse matrix in triplet form. 
- Constructor    
    - `new("dtTMatrix", ...)`  
    - `sparseMatrix(*, triangular=TRUE, giveCsparse=FALSE)`  
- Coercion
    - `as(*, "triangularMatrix")`
    - `as(*, "dtTMatrix")`  
- Examples   
```{r dtTMatrix}
(t1 <- new("dtTMatrix", x= c(3,7), 
          i= 0:1, j=3:2, Dim= as.integer(c(4,4))))

# Output
# 4 x 4 sparse Matrix of class "dtTMatrix"
#             
# [1,] . . . 3
# [2,] . . 7 .
# [3,] . . . .
# [4,] . . . .

i <- c(1,3:8); j <- c(2,9,6:10); x <- 7 * (1:7)
(tA <- sparseMatrix(i, j, x = x, 
                    triangular= TRUE, giveCsparse = FALSE))

# Output
# 10 x 10 sparse Matrix of class "dtTMatrix"
#                               
#  [1,] . 7 . . .  .  .  .  .  .
#  [2,] . . . . .  .  .  .  .  .
#  [3,] . . . . .  .  .  . 14  .
#  [4,] . . . . . 21  .  .  .  .
#  [5,] . . . . .  . 28  .  .  .
#  [6,] . . . . .  .  . 35  .  .
#  [7,] . . . . .  .  .  . 42  .
#  [8,] . . . . .  .  .  .  . 49
#  [9,] . . . . .  .  .  .  .  .
# [10,] . . . . .  .  .  .  .  .

mtxt <- c("1 0 0 3",
          "0 1 7 0",
          "0 0 1 0",
          "0 0 0 1")
M <- as.matrix(read.table(text=mtxt))
dimnames(M) <- NULL
dgT <- as(M, "dgTMatrix")

(dtT <- as(dgT, "triangularMatrix"))

# Output
# 4 x 4 sparse Matrix of class "dtTMatrix"
#             
# [1,] 1 . . 3
# [2,] . 1 7 .
# [3,] . . 1 .
# [4,] . . . 1

(dtT <- as(M, "dtTMatrix"))

# Output
# 4 x 4 sparse Matrix of class "dtTMatrix"
#             
# [1,] 1 . . 3
# [2,] . 1 7 .
# [3,] . . 1 .
# [4,] . . . 1
```

#### dsTMatrix {#dst}
- Description: symmetric numeric sparse matrix in triplet form. 
- Constructor    
    - `new("dsTMatrix", ...)`  
    - `sparseMatrix(*, symmetric=TRUE, giveCsparse=FALSE)`  
- Coercion    
    - `as(*, "symmetricMatrix")`
    - `as(*, "dsTMatrix")`  
- Examples   
```{r dsTMatrix}
(dsT <- new("dsTMatrix", i=as.integer(c(0,1,0,2,1,3,0,2,4)),
           j=as.integer(c(0,1,2,2,3,3,4,4,4)),
           x=as.double(c(10,10,1,10,1,10,3,1,10)),
           Dim=as.integer(c(5, 5))))

# Output
# 5 x 5 sparse Matrix of class "dsTMatrix"
#                    
# [1,] 10  .  1  .  3
# [2,]  . 10  .  1  .
# [3,]  1  . 10  .  1
# [4,]  .  1  . 10  .
# [5,]  3  .  1  . 10

i <- c(1,3:8); j <- c(2,9,6:10); x <- 7 * (1:7)
(sA <- sparseMatrix(i, j, x = x, 
                    symmetric = TRUE, giveCsparse = FALSE))

# Output
# 10 x 10 sparse Matrix of class "dsTMatrix"
#                                  
#  [1,] . 7  .  .  .  .  .  .  .  .
#  [2,] 7 .  .  .  .  .  .  .  .  .
#  [3,] . .  .  .  .  .  .  . 14  .
#  [4,] . .  .  .  . 21  .  .  .  .
#  [5,] . .  .  .  .  . 28  .  .  .
#  [6,] . .  . 21  .  .  . 35  .  .
#  [7,] . .  .  . 28  .  .  . 42  .
#  [8,] . .  .  .  . 35  .  .  . 49
#  [9,] . . 14  .  .  . 42  .  .  .
# [10,] . .  .  .  .  .  . 49  .  .

mm <- Matrix(toeplitz(c(10, 0, 1, 0, 3)), sparse = TRUE)
mT <- as(mm, "dgTMatrix")
(symM <- as(mT, "symmetricMatrix"))

# Output
# 5 x 5 sparse Matrix of class "dsTMatrix"
#                    
# [1,] 10  .  1  .  3
# [2,]  . 10  .  1  .
# [3,]  1  . 10  .  1
# [4,]  .  1  . 10  .
# [5,]  3  .  1  . 10

(dsT <- as(mT, "dsTMatrix"))

# Output
# 5 x 5 sparse Matrix of class "dsTMatrix"
#                    
# [1,] 10  .  1  .  3
# [2,]  . 10  .  1  .
# [3,]  1  . 10  .  1
# [4,]  .  1  . 10  .
# [5,]  3  .  1  . 10

```

#### dgRMatrix {#dgr}
- Description: general row-oriented numeric sparse matrix. 
- Constructor    
    - `new("dgRMatrix", ...)`  
- Coercion    
    - `as(*, "dgRatrix")`  
- Examples  
```{r dgRMatrix}
(dgR <- new("dgRMatrix", j=as.integer(c(0,2,1,3)),
            p=as.integer(c(0,2,3,3,4)),
            x=as.double(c(3,1,2,1)),
            Dim=as.integer(rep(4,2))))

# Output
# 4 x 4 sparse Matrix of class "dgRMatrix"
#             
# [1,] 3 . 1 .
# [2,] . 2 . .
# [3,] . . . .
# [4,] . . . 1

m <- matrix(c(3,0,0,0,0,2,0,0,1,0,0,0,0,0,0,1), nrow=4)
(dgR <- as(m, "dgRMatrix"))

# Output
# 4 x 4 sparse Matrix of class "dgRMatrix"
#             
# [1,] 3 . 1 .
# [2,] . 2 . .
# [3,] . . . .
# [4,] . . . 1
```

#### dtRMatrix {#dtr}
- Description: triangular row-oriented numeric sparse matrix. 
- Constructor    
    - `new("dtRMatrix", ...)`  
- Examples  
```{r dtRMatrix}
(m2 <- new("dtRMatrix", Dim = c(2L,2L),
           x = c(5, 1:2), p = c(0L,2:3), j= c(0:1,1L)))

# Output
# 2 x 2 sparse Matrix of class "dtRMatrix"
#         
# [1,] 5 1
# [2,] . 2
```

#### dsRMatrix {#dsr}
- Description: symmetric row-oriented numeric sparse matrix. 
- Constructor    
    - `new("dsRMatrix", ...)`  
- Coercion    
    - `as("dsCMatrix", "dsRMatrix")`  
- Examples  
```{r dsRMatrix}
(m2 <- new("dsRMatrix", Dim = c(2L,2L),
                  x = c(3,1), j = c(1L,1L), p = 0:2))

# Output
# 2 x 2 sparse Matrix of class "dsRMatrix"
#         
# [1,] . 3
# [2,] 3 1

m <- matrix(c(0, 3, 3, 1), nrow = 2)
dsC <- as(m, "dsCMatrix")
(dsR <- as(dsC, "dsRMatrix"))

# Output
# 2 x 2 sparse Matrix of class "dsRMatrix"
#         
# [1,] . 3
# [2,] 3 1
```

#### indMatrix {#ind}
- Description: index matrix.  
- Constructor    
    - `new("indMatrix", ...)`  
- Coercion    
    - `as(*, "indMatrix")`  
- Examples  
```{r indMatrix}
(ind <- new("indMatrix", perm = 2:4, Dim = 3:4))

# Output
# 3 x 4 sparse Matrix of class "indMatrix"
#             
# [1,] . | . .
# [2,] . . | .
# [3,] . . . |

(ind <- as(2:4, "indMatrix"))

# Output
# 3 x 4 sparse Matrix of class "indMatrix"
#             
# [1,] . | . .
# [2,] . . | .
# [3,] . . . |
```

#### pMatrix {#pm}
- Description: permutation matrix. 
- Constructor    
    - `new("pMatrix", ...)`  
- Coercion    
    - `as(*, "pMatrix")`  
- Examples  
```{r pMatrix}
(p <- new("pMatrix", perm = c(2L, 3L, 1L), Dim = rep(3L, 2)))

# Output
# 3 x 3 sparse Matrix of class "pMatrix"
#           
# [1,] . | .
# [2,] . . |
# [3,] | . .

(p1 <- as(c(2,3,1), "pMatrix"))

# Output
# 3 x 3 sparse Matrix of class "pMatrix"
#           
# [1,] . | .
# [2,] . . |
# [3,] | . .
```

#### ddiMatrix {#ddi}
- Description: numeric diagonal Matrix.  
- Constructor    
    - `new("ddiMatrix", ...)`
    - `Diagonal(*)`
- Examples  
```{r ddiMatrix}
(ddi <- new("ddiMatrix", diag = "U"))

# Output
# 5 x 5 diagonal matrix of class "ddiMatrix"
#      [,1] [,2] [,3] [,4] [,5]
# [1,]    1    .    .    .    .
# [2,]    .    1    .    .    .
# [3,]    .    .    1    .    .
# [4,]    .    .    .    1    .
# [5,]    .    .    .    .    1

(ddi <- new("ddiMatrix", x = 10*(1:5), Dim = rep(5L, 2)))

# Output
# 5 x 5 diagonal matrix of class "ddiMatrix"
#      [,1] [,2] [,3] [,4] [,5]
# [1,]   10    .    .    .    .
# [2,]    .   20    .    .    .
# [3,]    .    .   30    .    .
# [4,]    .    .    .   40    .
# [5,]    .    .    .    .   50


(I5 <- Diagonal(5))

# Output
# 5 x 5 diagonal matrix of class "ddiMatrix"
#      [,1] [,2] [,3] [,4] [,5]
# [1,]    1    .    .    .    .
# [2,]    .    1    .    .    .
# [3,]    .    .    1    .    .
# [4,]    .    .    .    1    .
# [5,]    .    .    .    .    1

(D5 <- Diagonal(x = 10*(1:5)))

# Output
# 5 x 5 diagonal matrix of class "ddiMatrix"
#      [,1] [,2] [,3] [,4] [,5]
# [1,]   10    .    .    .    .
# [2,]    .   20    .    .    .
# [3,]    .    .   30    .    .
# [4,]    .    .    .   40    .
# [5,]    .    .    .    .   50
```

### Sparse Matrix in "Armadillo" {#smia}  
In the library "Armadillo", sparse matrix is currently stored as [CSC](https://en.wikipedia.org/wiki/Sparse_matrix#Compressed_sparse_column_.28CSC_or_CCS.29) format. Such kind of format is quite similar to numeric column-oriented sparse matrix in the library "Matrix"(including dgCMatrix, dtCMatrix and dsCMatrix). But provided the index in C++ starts from 0, the element access should be specially paid attention to when RcppArmadillo is applied.

## Member functions {##member-funcs}
### Element Access {#elem-access}
- Description: access the element of a sparse matrix (the index should start from 0).
- Functions
    - `(n)`: access the n-th element under the assumption of a flat layout, with column-major ordering of data and a bounds check.  
    - `.at(n)` or `[n]`: similar to `(n)` without a bounds check.  
    - `(i,j)`:  access the element in i-th row and j-th column with a bounds check.  
    - `.at(i,j)`:  similar to `(i, j)` without a bounds check.  
- Examples  
```{Rcpp elem_access}
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

using namespace Rcpp;
using namespace arma;

//[[Rcpp::export]]
double elem(sp_mat X, int row_number, int col_number) {
  return X(row_number, col_number); 
}

/*** R
library(Matrix)
dgC <- Matrix(c(0,0,2:0), 3,5)
elem(dgC, 0, 1)
*/

# Output
# [1] 1

```

### Operators {#operations}
- Operators   
    - `+`: Addition of two matrices.  
    - `-`: Subtract of a matrx from another.    
    - `*`: Multiplication of two matrices.  
    - `/`: Element-wise division of a matrix by another.    
    - `%`: Element-wise multiplication of two matrices.    
    - `==`: Elevaluate if two matrices are equal.  
    - `!=`: Elevaluate if two matrices are not equal.   
    - `<=`: Elevaluate if a matrix is less than or equal to another.  
    - `>=`: Elevaluate if a matrix is greater than or equal to another.    
    - `<`: Elevaluate if a matrix is less than another.  
    - `>`: Elevaluate if a matrix is greater than another.  
- Examples  
```{Rcpp add&multiply}
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

using namespace Rcpp;
using namespace arma;

//[[Rcpp::export]]
sp_mat add(sp_mat A, sp_mat B) {
  return A + B;
}

sp_mat multiply(sp_mat A, sp_mat B) {
  return A * B;
}

/*** R

*/

# Output

```

### Submatrix Views {#sub-views}
- Description: Access the row, the column or the submatrix of a sparse matrix.  
- Functions  
    - `X.col(j)`: access the j-th column.  
    - `X.row(i)`: access the i-th row.  
    - `X.cols(first_col, last_col)`: access the columns starting from first_col to last_col.    
    - `X.rows(first_row, last_row)`: access the rows starting from first_row to last_row.  
    - `X.submat(first_row, first_col, last_row, last_col)`: access the submatrix starting from first_row and first_col to last_row and last_col.  
    - `X(span(first_row, last_row), span(first_col, last_col))`: access the submatrix starting from first_row and first_col to last_row and last_col.    
    - `X(first_row, first_col, size(n_rows, n_cols))`: access the submatrix starting from first_row and first_col with size of n_rows and n_cols.  
    - `X(first_row, first_col, size(Y))`: access the submatrix starting from first_row and first_col with size of another matrix Y.  
    - `X.head_cols(number_of_cols)`: access the head number_of_cols columns.  
    - `X.head_rows(number_of_rows)`: access the head number_of_rows rows.   
    - `X.tail_cols(number_of_cols)`: access the tail number_of_cols columns.  
    - `X.tail_rows(number_of_rows)`: access the tail number_of_rows rows.  
    - `X.unsafe_col(j)`: access the j-th column.   
    
- Examples
```{Rcpp sub-views}
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

using namespace Rcpp;
using namespace arma;

//[[Rcpp::export]]
vec col(sp_mat X, int col_number) {
  return X.col(col_number);
}

//[[Rcpp::export]]
vec row(sp_mat X, int row_number) {
  return X.row(row_number);
}

//[[Rcpp::export]]
sp_mat cols(sp_mat X, int first_col, int last_col) {
  return X.cols(first_col, last_col);
}

//[[Rcpp::export]]
sp_mat rows(sp_mat X, int first_row, int last_row) {
  return X.rows(first_row, last_row);
}

//[[Rcpp::export]]
sp_mat submat(sp_mat X, int first_row, int first_col, int last_row, int last_col) {
  return X.submat(first_row, first_col, last_row, last_col);
}

//[[Rcpp:export]]
sp_mat head_cols(sp_mat X, int number_of_cols) {
  return X.head_cols(number_of_cols);
}

//[[Rcpp::export]]
sp_mat head_rows(sp_mat X, int number_of_rows) {
  return X.head_rows(number_of_rows);
}

//[[Rcpp::export]]
sp_mat tail_cols(sp_mat X, int number_of_cols) {
  return X.tail_cols(number_of_cols);
}
```

### Diagonal Views {#diag-views}
- Description: access the diagonal value of a sparse matrix. 
- Function
    - `.diag()`: access the main diagonal of a sparse matrix.  
    - `.diag(k)`:  
          - When k = 0, access the main diagonal of a sparse matrix.  
          - When k > 0, access the k-th super-diagonal of a sparse matrix.   
          - When k < 0, access the k-th sub-diagonal of a sparse matrix.  
- Examples
```{Rcpp diag-views}
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

using namespace Rcpp;
using namespace arma;

//[[Rcpp::export]]
vec diag(sp_mat X, int k) {
  return X.diag(k);
}
```

### Savings & Loadings {#save-load}
- Description:
- Function
    - `.save(name)`
    - `.save(name, file_type)`
    - `.save(stream)`
    - `.save(stream, file_type)`
    - `.load(name)`
    - `.load(name, file_type)`
    - `.load(stream)`
    - `.load(stream, file_type)`
- Examples
```{Rcpp save-load}
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

using namespace Rcpp;
using namespace arma;

//[[Rcpp::export]]
void save(sp_mat X, std::string name) {
  X.save(name);
}

//[[Rcpp::export]]
sp_mat load(sp_mat X, std::string name) {
  X.load(name); 
  return X;
}
```

### Element-wise Functions {#elem-wise-funcs}
- Description: functions that undertakes element-wise operations.
- Functions
    - `abs(X)`: obtain the absolute value of each element.
    - `sqrt(X)`: obtain the square root of each element.
    - `square(X)`: obtain the squared value of each element.
    - `floor(X)`: (To be continued)
    - `ceil(X)`: (To be continued)
    - `round(X)`: round to nearest integer, with halfway cases rounded away from zero.  
    - `trunc(X)`: round to nearest integer, towards zero.  
    - `sign(X)`: sign function.
- Examples
```{Rcpp abs}
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

using namespace Rcpp;
using namespace arma;

//[[Rcpp::export]]
sp_mat abs(sp_mat X) {
  return abs(X);
}

//[[Rcpp::export]]
sp_mat sqrt(sp_mat X) {
  return sqrt(X);
}

//[[Rcpp::export]]
sp_mat square(sp_mat X) {
  return square(X);
}

//[[Rcpp::export]]
sp_mat floor(sp_mat X) {
  return floor(X);
}

//[[Rcpp::export]]
sp_mat ceil(sp_mat X) {
  return ceil(X);
}

//[[Rcpp::export]]
sp_mat round(sp_mat X) {
  return round(X);
}

//[[Rcpp::export]]
sp_mat trunc(sp_mat X) {
  return trunc(X);
}

//[[Rcpp::export]]
sp_mat sign(sp_mat X) {
  return sign(X);
}
```

### Scalar Functions {#scalar-funcs}
- Description: include functions that generate a scalar.  
- Functions  
    - `accu(X)`: accumulate all elements of a sparse matrix X.   
    - `dot(A, B)`: obtain the dot product of sparse matrices A and B.  
    - `norm(X, p)`: obtain the p-norm of a sparse matrix X. p can be one of 1, 2, "inf", "fro" and can also be omitted, in which p is 2 by default.  
    - `trace(X)`: obtain the trace of a sparse matrix.  
- Examples  
```{Rcpp accu}
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

using namespace Rcpp;
using namespace arma;

//[[Rcpp::export]]
double accu(sp_mat X) {
  return accu(X);
}

//[[Rcpp::export]]
double dot(sp_mat A, sp_mat B) {
  return dot(A, B);
}

//[[Rcpp::export]]
double norm(sp_mat X, int p) {
  return norm(X, p);
}

//[[Rcpp::export]]
double trace(sp_mat X) {
  return trace(X);
}
```

### Vector Valued Functions {#vec-funcs}
- Description: include functions that generate a vector.   
- Functions  
    - `min(X, dim)`: obtain the maximum value of each column when dim = 0, or each row when dim = 1. The dim is optional; by default dim = 0 is used.  
    - `max(X, dim)`: obtain the minimum value of each column when dim = 0, or each row when dim = 1. The dim is optional; by default dim = 0 is used.     
    - `nonzeros(X)`: obtain the non-zero values of a sparse matrix X.  
    - `sum(X, dim)`: obtain the sum of all elements in each column when dim = 0, or each row when dim = 1. The dim is optional; by default dim = 0 is used.  
    - `mean(X, dim)`: obtain the mean of all elements in each column when dim = 0, or each row when dim = 1. The dim is optional; by default dim = 0 is used.  
    - `var(X, dim)`: obtain the variance of all elements in each column when dim = 0, or each row when dim = 1. The dim is optional; by default dim = 0 is used.  
- Examples  
```{Rcpp max}
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

using namespace Rcpp;
using namespace arma;

//[[Rcpp::export]]
vec max(sp_mat X, int dim) {
  return max(X, dim);
}

//[[Rcpp::export]]
vec min(sp_mat X, int dim) {
  return min(X, dim);
}

//[[Rcpp::export]]
vec nonzeros(sp_mat X) {
  return nonzeros(X);
}

//[[Rcpp::export]]
vec sum(sp_mat X, int dim) {
  return sum(X, dim);
}

//[[Rcpp::export]]
vec mean(sp_mat X, int dim) {
  return mean(X, dim);
}

//[[Rcpp::export]]
vec var(sp_mat X, int dim) {
  return var(X, dim);
}
```

### Matrix Valued Functions {#mat-funcs}
- Description: include functions that generate a matrix.  
- Functions  
    - `diagmat(X, k)`: create a matrix with the k-th diagonal containing a copy of the k-th diagonal of X; all other elements are set to zeros. The k is optional; by default k = 0 is used.   
    - `trimatu(X, k)`: create a matrix with the upper triangular part above k-th diagonal containing a copy of the upper triangular part above k-th diagonal of X; all other elements are set to zeros. The k is optional; by default k = 0 is used.   
    - `trimatl(X, k)`: create a matrix with the lower triangular part below k-th diagonal containing a copy of the lower triangular part above k-th diagonal of X; all other elements are set to zeros. The k is optional; by default k = 0 is used.  
    - `symmatu(X)`: create a symmetric matrix by reflecting the upper triangular part to the lower triangular one.  
    - `symmatl(X)`: create a symmetric matrix by reflecting the lower triangular part to the upper triangular one.  
    - `join_rows(A, B)`: horizontal concatenation.   
    - `join_cols(A, B)`: vertical concatenation.  
    - `repmat(X, num_copies_per_row, num_copies_per_col)`: create a matrix by replicating the matrix with each row copied num_copies_per_row times and each column copied num_copies_per_col times.  
    - `reshape(X, n_rows, n_cols)`: create a n_rows by n_cols matrix whose elements are from X in a column-oriented way.  
    - `reshape(X, size(X))`: similar to `reshape(X, n_rows, n_cols)`, but the size of the new matrix is size of the matrix X.  
    - `resize(X, n_rows, n_cols)`: create a n_rows by n_cols matrix whose elements are from X in a column-oriented way.  
    - `resize(X, size(X))`: similar to `reshape(X, n_rows, n_cols)`, but the size of the new matrix is size of the matrix X.
    - `.t()`: create a transposed copy of a matrix.  
    - `trans(X)`: create a transposed copy of a matrix X.  
- Examples  
```{Rcpp diagmat}
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

using namespace Rcpp;
using namespace arma;

//[[Rcpp::export]]
sp_mat diagmat(sp_mat X) {
  return diagmat(X);
}

//[[Rcpp::export]]
sp_mat trimatu(sp_mat X) {
  return trimatu(X);
}

//[[Rcpp::export]]
sp_mat trimatl(sp_mat X) {
  return trimatl(X);
}

//[[Rcpp::export]]
sp_mat symmatu(sp_mat X) {
  return symmatu(X);
}

//[[Rcpp::export]]
sp_mat symmatl(sp_mat X) {
  return symmatl(X);
}

//[[Rcpp::export]]
sp_mat join_rows(sp_mat A, sp_mat B) {
  return join_rows(A, B);
}

//[[Rcpp::export]]
sp_mat join_cols(sp_mat A, sp_mat B) {
  return join_cols(A, B);
}

//[[Rcpp::export]]
sp_mat repmat(sp_mat A, int num_copies_per_row, int num_copies_per_col) {
  return repmat(A, num_copies_per_row, num_copies_per_col);
}

//[[Rcpp::export]]
sp_mat reshape(sp_mat A, int n_rows, int n_cols) {
  return reshape(A, n_rows, n_cols);
}

//[[Rcpp::export]]
sp_mat resize(sp_mat A, int n_rows, int n_cols) {
  return resize(A, n_rows, n_cols);
}

//[[Rcpp::export]]
sp_mat trans(sp_mat A) {
  return trans(A);
}
```

### Generated Matrices {#generated-mats}
- Description: create a new sparse matrix based on different specifications.  
- Functions  
    - `speye(n_rows, n_cols)`: create a n_rows by n_cols sparse matrix with unit values in main diagonal and zeros off the diagonal.  
    - `speye(size(X))`: similar to `speye(n_rows, n_cols)`, but the size is same as that of the matrix X.  
    - `spones(X)`: create a sparse matrix similar to the sparse matrix X, but with non-zero elements set to one.  
    - `sprandu(n_rows, n_cols, density)`: create a n_rows by n_cols sparse matrix with non-zero elements drawn from a uniform distribution in the [0, 1] interval.  
    - `sprandu(size(X), density)`: similar to `sprandu(n_rows, n_cols, density)`, but the size of the new matrix is the same as the matrix X.  
    - `sprandn(n_rows, n_cols, density)`: create a n_rows by n_cols sparse matrix with non-zero elements drawn from a standard normal distribution.  
    - `sprandn(size(X), density)`: similar to `sprandu(n_rows, n_cols, density)`, but the size of the new matrix is the same as the matrix X.  
- Examples  
```{Rcpp speye}
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

using namespace Rcpp;
using namespace arma;

//[[Rcpp::export]]
sp_mat speye(int n_rows, int n_cols) {
  return speye(n_rows, n_cols);
}

//[[Rcpp::export]]
sp_mat spones(sp_mat A) {
  return spones(A);
}

//[[Rcpp::export]]
sp_mat sprandu(int n_rows, int n_cols, double density) {
  return sprandu(n_rows, n_cols, density);
}

//[[Rcpp::export]]
sp_mat sprandn(int n_rows, int n_cols, double density) {
  return sprandn(n_rows, n_cols, density);
}
```

### Eigen & Singular Value Decomposition {#evd-svd}
- Description:  
- Functions  
    - `eigs_sym()`  
    - `eigs_gen()`  
    - `svds()`  
- Examples
### Solution of Sparse Linear Systems {#solve}
- Description:  
- Functions  
    - `spsolve()`  
- Examples  
### Miscellaneous {#misc}  

